---
title: "Brydes Data Analysis"
output: html_document
---

```{r prep results="hide"}
#| echo: false
#| warnings: false
library(here)
here()
source(here::here("_code/brydes/_brydesCommonR.R"))

#Only run if you NEED to rerun the raw data for deploy, review, raven, raven_effort fields
#source(here::here("_code/brydes/rawDataPrep_brydes.R")) 

```
# Manual Review and Annotation
Manual review was performed by an experienced researcher (SR) using RavenPro 1.6. For each recording with usable data, measurements were made to a selection table (see SWFSC-PAM-Sonobuoy-RavenPro-MeasurementPreset.ravenpreset in supplement folder in this repo) and three custom annotation columns were included (Start_End, Detection, Quality, following the SWFSC_PAM_Sonobuoy_Annotation_keymap.ravenpreset in the supplement folder in this repo).

Effort was identified by a start effort (Start_End, Detection, Quality annotations of 1, 9, 9) at the beginning of the usable recording, and end of effort was identified (3, 9, 9) at the end of the usable effort. Bouts of noise were identified within RavenPro by a temporary end of the effort due to noise (end_noise, 3, 1, 9), and a restart after a bout of noise was annotated (start_noise, 1, 9, 9). Detections of potential biological sounds were identifed as a detection of a specific species (where Start_End = 9, Detection = species, and Quality = low/med/high). Selection tables were saved using the following naming format: Cruise#_RecordingID_YYMMDD_HHmm.Table1.selections.txt.

For mid-frequency review, recordings decimated to 4800 Hz were reviewed using RavenPro Software (v1.6.4). Page size = 60 seconds, 90% increment, 10% step increment. View only spectrogram (not waveform), 2 rows with 512 FFT with 200 Hz to 2400 Hz in window. For each recording with usable data, measurements were made to a selection table (see SWFSC-PAM-Sonobuoy-RavenPro-MeasurementPreset.ravenpreset in supplement folder in this repo) and three custom annotation columns were included (Start_End, Detection, Quality, following the SWFSC_PAM_Sonobuoy_Annotation_keymap.ravenpreset in the supplement folder in this repo).

For low-frequency review....[have not yet done low frequency review]


# Data Prep
Load-Merge-Modify Data

To Do: offsetGMT (add to this data, add UTC column, and develop a testing script to see if there are deeper issues with swfscDAS or the raw das data)


```{r prepData}
deploy <- readRDS(here("data/brydes/deploy.RDS"))

raven <- readRDS(here("data/brydes/raven.RDS"))

raven_effort <- readRDS(here("data/brydes/raven_effort.RDS"))

review <- readRDS(here("data/brydes/review.RDS")) %>%
    filter(!(cruise == 1623 & dat_number %in% c(21, 24, 25)))%>% #too much interference from Mn song
    filter(!(cruise == 1635 & dat_number == 40)) #This data is missing from the deploy, so can't use!
# Add count of Annotations to Review
annotation_col <- "be500"
review <- count_true_annotations(review, raven, annotation_col)
#add raven_effort
review <- merge_deploy_review(review, raven_effort, remove_unmatched_deploy = FALSE)
## Note
# Warning in merge_deploy_review(review, raven_effort, remove_unmatched_deploy = FALSE) :
#   Found 4 row(s) in 'review' that do not exist in 'deploy'!
# Showing the first few unmatched pairs:
# # A tibble: 4 × 2
#   cruise dat_number
#   <chr>  <chr>     
# 1 1623   6         
# 2 1635   18        
# 3 1635   22        
# 4 1635   40



# Flag any review_effort rows with no match in deploy
unmatched <- review %>%
  anti_join(deploy, by = c("cruise", "dat_number"))

if (nrow(unmatched) > 0) {
  message("WARNING: ", nrow(unmatched), " row(s) in review_effort have no match in deploy:")
  unmatched %>%
    select(cruise, dat_number) %>%
    print()
}
# Each of these, below, were additional recordings (tapes) for a sonobuoy/site
# WARNING: 7 row(s) in review_effort have no match in deploy:
# # A tibble: 7 × 2
#   cruise dat_number
#   <chr>  <chr>     
# 1 1621   4         
# 2 1621   9         
# 3 1621   13        
# 4 1629   5         
# 5 1629   6         
# 6 1629   8         
# 7 1631   35  
```
## RETURN TO FIX THIS!! ##

``` {r testTime}
source(here("_code/brydes/compare_dates.R"))
result <- compare_dates(deploy, survey_data)
result

```
#OffsetGMT
The times provided in the raw datasets are 'local time', which is not related to longitude/time zones, but rather they are a 'floating' local time to the ship. The ship would shift the clock (effectively change time zones) according to sun/meal times. These times were also not always changed immediately (there are notes in some das files, 'oops, forgot to change the time'). Also, the GMT offset saved in the original DAS files is inverted, and it is extra confusing near the date line (HICEAS surveys, at the least). The swfscDAS R package has a script to identify the OffsetGMT for each line, but there are some issues where it  may not be calculated (eg, near the date line). See here for more info: https://swfsc.github.io/swfscDAS/articles/swfscDAS.html#read-and-process-data.

To examine diurnal variation in calling behavior, I need to convert times to GMT/UTC and then look at sun angle for that UTC at that Lat/Long. Rather than convert ALL of the data (for which there will be errors), I will instead Read & Process the data and search for the 24 hours surrounding a specific detection to ensure the OffsetGMT is stable. If yes, then I will import that as date_UTC into my dataframe. If not-- then I will want to manually review to see if there is a problem.  

```{r gmtOffset}
# =============================================================================
# USAGE
# =============================================================================

# Step 1 — adds offsetGMT; review CHANGE rows before proceeding
deploy <- brydes_df %>% add_offset_gmt()

# Optional: inspect rows needing manual review
brydes_df %>% filter(offsetGMT == "CHANGE" | is.na(offsetGMT)) %>% View()
# 1623-2 is ok at 7
# 1623-4 is ok at 6
# 1623-5 is ok at 6
# 1624-8 is ok at 8
# 1631-2 is ok at 9
# 1631-3 is ok at 9
# 1631-4 is ok at 9
# 1631-36 is ok at 5
# 1631-37 is ok at 6  #Check that the date on this is 10/19/2006!!! When I rerun things
# 1631-44 is ok at 8


# Step 2 — convert to UTC once you're satisfied with offsetGMT values
brydes_df <- brydes_df%>% add_date_utc()

```



# Detection Map

```{r map results='hide'}
source(here::here("_code/brydes/detection_map.R"))
detection_df <- merge_deploy_review(deploy, review, remove_unmatched_deploy = TRUE)
#Note: 7 rows in 'review' that do not exist in 'deploy', because those had multiple reviewed tapes per deployed buoy

detection_map(
  detection_df,
  output_name = "brydes/brydes_map",
  species = list(
    `Any Bryde's call` = list(
      detected = "be",
      sighting = "72_sightings",
      color    = "#FB8C00",
      label    = "Bryde's MF Calls"
    ),
    `500 Hz calls` = list(
      detected  = "be500",
      sighting  = "72_sightings",
      threshold = 0,
      color     = "#FFF176",
      label     = "500 Hz calls"
    )
  ),
  show_not_detected = TRUE,
  padding_deg        = 4,
  legend_title      = "Recordings",
  label_sighting     = "Bryde's sighting",
  label_no_sighting  = "Acoustic only",
  label_not_detected = "No Acoustic Detections"
)

```



# IPI Histogram

```{r ipiHist}
source(here::here("_code/brydes/analyze_burst_ipi.R"))
be500 <- raven %>%
  filter(be500 == "Y")
# Auto peak detection (default)
ipi_result <- analyze_burst_ipi(be500, maxIPI = 60.0, bin_width = 1, show_median = FALSE, 
                            table_path = here("output/brydes/brydes_ipi_summary_table.docx"),
                            plot_path = here("output/brydes/brydes_ipi_histogram.png"))
ipi_result


```


#Plot Detection Values

```{r be500}
be500 <- raven%>%
  filter(be500 == "Y")
be500_plot <- create_acoustic_violin_plots(be500)
be500_plot

ggplot2::ggsave(
  filename = here("output/brydes/be500_violin.png"),
  plot = be500_plot,
  width = 10,       # Width in inches
  height = 6,       # Height in inches
  dpi = 300,        # Standard academic publication resolution
  bg = "white"      # Explicitly keeps background white (prevents transparent artifacts)
)


# Example: Pushes the labels further to the right (hjust = -0.5) 
# and higher above the line (vjust = -1)
# create_acoustic_violin_plots(be500_raven, hjust = -0.5, vjust = -1)


```

# Diurnal Plot
```{r diurnalPlot}
source(here("_code/brydes/diurnal_polar_plot.R"))

diurnalData <- merge_deploy_review(deploy, review, remove_unmatched_deploy = TRUE)
# diurnalData$date <- as.POSIXct(diurnalData$date, format = "%Y-%m-%d %H:%M:%S", tz = "UTC")

#Note: 7 rows in 'review' that do not exist in 'deploy', because those had multiple reviewed tapes per deployed buoy

plot_diurnal_calls(
  diurnalData,
  be_color    = "#FB8C00",
  be500_color = "#FFF176",
  day_color   = "white",
  dawn_color  = "#F6F6FF",
  night_color = "gray",
  shared_effort_scale = TRUE,
  plot_title  = NULL,
  be_title    = "Bryde's Mid-Frequency Calls",
  be500_title = "Bryde's 500-Hz Calls",
  save_path   = here("output/brydes/diurnal_plot.png"),
  width = 14, height = 7,
  force_dawn = 5, force_sunrise = 6.0,
  force_sunset = 18.0, force_dusk = 19,
  show_det_labels = FALSE
)



# If your columns have different names
#plot_diurnal_calls(diurnalData, date_col = "start_time", lat_col = "lat", lon_col = "lon")
```


# Table of deployments, good recording time, species deployed on, species within 10nmi or 2 hours

Cruise name, deploy ID, date/time, Sighting # & Species, Species w/in 10nmi or 2 hours
```{r deployTable}
deployTableData <- merge_deploy_review(deploy, review, remove_unmatched_deploy = TRUE)
#Note: 7 rows in 'review' that do not exist in 'deploy', because those had multiple reviewed tapes per deployed buoy

# 1. Prepare the data for the table
table_data <- deployTableData %>%
  dplyr::mutate(`Sonobuoy ID` = paste(cruise, dat_number, sep = "-")) %>%
  dplyr::select(
    `Sonobuoy ID`, 
    date, latitude, longitude, "72_sightings", 
    effort_min, recording_quality_code, be, be500
  ) %>%
  mutate(recording_quality_code = str_to_sentence(recording_quality_code))%>%
  mutate(latitude = as.numeric(latitude),
         longitude = as.numeric(longitude))%>%
  dplyr::rename_with(~ stringr::str_to_title(stringr::str_replace_all(.x, "_", " ")))

#Choose which table to run
 table_data <- table_data %>%
   filter(Be %in% TRUE)
 
 # table_data <- table_data %>%
 #   filter(!Be %in% TRUE)

# 3. Build the wide publication-ready table
publication_table <- table_data %>%
  gt() %>%
  cols_label(
    `72 Sightings` = "Bryde's Sightings",
    'Be' = "Bryde's MF Calls",
    'Be500' = "Bryde's 500 Hz"
  ) %>%
  # Format column headers to be bold
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  # Change decimal places for columns
    fmt_number(
    columns = `Effort Min`,
    decimals = 1
  )%>%
  fmt_number(
    columns = c(Latitude, Longitude),
    decimals = 3
  )%>%
  # Add a double line beneath the column headers
  tab_style(
    style = cell_borders(
      sides = "bottom",
      color = "black",
      style = "double",
      weight = px(3)
    ),
    locations = cells_column_labels()
  ) %>%
  # Make the table wide and prevent text wrapping
  tab_options(
    table.width = pct(100),                         # Forces table to occupy full width
    container.width = pct(100),                     # Ensures container expands as well
    table.border.top.color = "black",
    table.border.top.width = px(2),
    table.border.bottom.color = "black",
    table.border.bottom.width = px(2),
    column_labels.border.top.color = "transparent",
    data_row.padding = px(6)
  ) %>% 
  # Force all text to stay on a single line so columns expand to fit data width
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = list(cells_body(), cells_column_labels())
  )

# Preview the table
publication_table
#Choose Save table for Brydes Detections
gtsave(data = publication_table, filename = here("output/brydes/brydes_DeployTable.png"),zoom = 2,
       vwidth = 1600, vheight = 900,)
gtsave(publication_table, here("output/brydes/brydes_DeployTable.docx"))

#Save Supplementary table with OTHER detections
# gtsave(data = publication_table, filename = here("output/brydes/supplementary_DeployTable.png"),zoom = 2,
#        vwidth = 1600, vheight = 900,)
# gtsave(publication_table, here("output/brydes/supplementary_DeployTable.docx"))

```

#Table of detections
For each call type, give median and 10th, 90th percentiles for: duration of call (s), low frequency (Hz), high frequency (Hz), center frequency (Hz), peak frequency (Hz), delta frequency (Hz), and signal-to-noise ratio (dB). Where appropriate, give ICI and IPI.
```{r be500Table}
be500 <- raven %>%
  filter(be500 == "Y")%>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

ipi_summary <- analyze_burst_ipi(be500, maxIPI = 30)
  

# View your new summary dataframe
print(ipi_summary)

#Make the table
publication_table <- ipi_summary %>%
  gt() %>%
  # Clean up column names to look professional for publication
  cols_label(
    sonobuoy_ID = "Sonobuoy ID",
    exceeded_max_count = "Breaks (> maxIPI)",
    mean_IPI = "Mean IPI (s)",
    median_IPI = "Median IPI (s)",
    range_IPI = "IPI Range (s)"
  ) %>%
  # Format numeric columns to a standard 3 decimal places
  fmt_number(
    columns = c(mean_IPI, median_IPI),
    decimals = 1
  ) %>%
  # Align text beautifully (ID left-aligned, metrics centered)
  cols_align(
    align = "left",
    columns = sonobuoy_ID
  ) %>%
  cols_align(
    align = "center",
    columns = c(exceeded_max_count, mean_IPI, median_IPI, range_IPI)
  ) %>%
  # Format column headers to be bold
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  # Add a double line beneath the column headers
  tab_style(
    style = cell_borders(
      sides = "bottom",
      color = "black",
      style = "double",
      weight = px(3)
    ),
    locations = cells_column_labels()
  ) %>%
  # Make the table wide and prevent text wrapping
  tab_options(
    table.width = pct(100),                          # Forces table to occupy full width
    container.width = pct(100),                      # Ensures container expands as well
    table.border.top.color = "black",
    table.border.top.width = px(2),
    table.border.bottom.color = "black",
    table.border.bottom.width = px(2),
    column_labels.border.top.color = "transparent",
    data_row.padding = px(6)
  ) %>% 
  # Force all text to stay on a single line so columns expand to fit data width
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = list(cells_body(), cells_column_labels())
  )

# Print the table
publication_table

gtsave(data = publication_table, filename = here("output/brydes/brydes_IPITable.png"),zoom = 2, 
       vwidth = 1600, vheight = 900,)
gtsave(publication_table, here("output/brydes/brydes_IPITable.docx"))
```
