---
title: "Brydes Data Analysis"
output: html_document
---
Give explicit instructions for working with data in Raven


# Merge survey data and analysis to create a single sonobuoy data file (this should probably go into the _commonR.r file and saved??)
```{r surveyDataprep}
source(here::here("_code/_commonR.R"))
library(Rraven)
library(warbleR)

surveyData <- read_csv(here("_data/surveyData.csv")) %>%
  clean_names()

review <- read.csv(here("_data/swfsc_sonobuoy_manualReview.csv"))%>%
  clean_names() 
#Next, merge these so that all the files in review have the important info (date, time, lat long,etc). Might need to be sure there is a specific ID column like CruiseNum_SBnum or something

surveyBrydes <- xxx #subset this data to only include survey data where "Brydes" detections are TRUE
  


```

Import/Merge/Modify Selection Tables
```{r ravenDataprep}
# List all your files
myPath <- "W:/Sonobuoys/d10_SelectionTables"#path for selection tables

list.files(path = myPath , pattern = "\\.txt$")  # List files in myPath


# Import raven data (providing the name of the column with the sound file names) in format that accommodates warbleR functions. Note, if you want to work with the time series (in warblR), then add argument <waveform = TRUE>.
d10 <- imp_raven(sound.file.col = "End.File", all.data = TRUE, ext.case = 'upper',
                 name.from.file = TRUE, path = myPath) 
#to work with warbleR, use d10w
d10w <- imp_raven(sound.file.col = "End.File", all.data = TRUE, ext.case = 'upper',
                     name.from.file = TRUE, warbler.format =  TRUE, path = myPath) 


## If I want to extract time series for use in warbleR
# create spectrograms and put them in a folder for review


# can I use this to export wav files of selections with a buffer?

# spectro-analysis in warbleR, can also measure harmonic features

# cross-correlation-- pair-wise similarity of the sounds, might want to do to quantify sounds from different encounters. Gives a similarity matrix to see which sounds are more similar to each other.




```






# Brydes Map

Map with bathymetry 
- Small black dots for brydes whale detections 
- Small yellow dots for brydes whale detections with calls
- Small yellow stars for brydes whale detections with 500 Hz pulses 

# Table of deployments, good recording time, species deployed on, species within 10nmi or 2 hours

Cruise name, deploy ID, date/time, Sighting # & Species, Species w/in 10nmi or 2 hours
```{r deployTable}



```

 
#Table of detections
For each call type, give median and 10th, 90th percentiles for: duration of call (s), low frequency (Hz), high frequency (Hz), center frequency (Hz), peak frequency (Hz), delta frequency (Hz), and signal-to-noise ratio (dB). Where appropriate, give ICI and IPI.