---
title: "Brydes Data Analysis"
output: html_document
---

```{r prep}
library(here)
here()
source(here::here("_code/_brydesCommonR.R"))

```
# Manual Review and Annotation
Manual review was performed by an experienced researcher (SR) using RavenPro 1.6. For each recording with usable data, measurements were made to a selection table (see SWFSC-PAM-Sonobuoy-RavenPro-MeasurementPreset.ravenpreset in supplement folder in this repo) and three custom annotation columns were included (Start_End, Detection, Quality, following the SWFSC_PAM_Sonobuoy_Annotation_keymap.ravenpreset in the supplement folder in this repo).

Effort was identified by a start effort (Start_End, Detection, Quality annotations of 1, 9, 9) at the beginning of the usable recording, and end of effort was identified (3, 9, 9) at the end of the usable effort. Bouts of noise were identified within RavenPro by a temporary end of the effort due to noise (end_noise, 3, 1, 9), and a restart after a bout of noise was annotated (start_noise, 1, 9, 9). Detections of potential biological sounds were identifed as a detection of a specific species (where Start_End = 9, Detection = species, and Quality = low/med/high). Selection tables were saved using the following naming format: Cruise#_RecordingID_YYMMDD_HHmm.Table1.selections.txt.

For mid-frequency review, recordings decimated to 4800 Hz were reviewed using RavenPro Software (v1.6.4). Page size = 60 seconds, 90% increment, 10% step increment. View only spectrogram (not waveform), 2 rows with 512 FFT with 200 Hz to 2400 Hz in window. For each recording with usable data, measurements were made to a selection table (see SWFSC-PAM-Sonobuoy-RavenPro-MeasurementPreset.ravenpreset in supplement folder in this repo) and three custom annotation columns were included (Start_End, Detection, Quality, following the SWFSC_PAM_Sonobuoy_Annotation_keymap.ravenpreset in the supplement folder in this repo).

For low-frequency review....[have not yet done low frequency review]


# Summary Table Prep
Download google table and save as a dataframe for downstream processing

# Data Prep
## 1. Import-Merge-Modify Selection Tables

```{r surveyDataprep}

# TO Do:
### --> Review 1623 #21, 24, 25 (humpbacks rule the day, but need to review for be500, label other bp as 'unknown')
### --> For each sighting, identify the species included in that sighting (or nearby), and then create a column for 'be" to identify if any are present.
### --> save each of the relevant R objects as dataframes in the data/ folder so I can skip this when I"m done
### --> check all times to see if they are local, can I change them to UTC?? Note: I emailed Sam, but may need to use tmz=""
### --> Clean up the code chunks (insert as much as possible into the functions)

### sb_review
# sonobuoy deployments that I reviewed manually
be_review <- read.csv(here("data/swfsc_sonobuoy_manualReview.csv"))%>%
  clean_names() %>%
  rename(sonobuoy_number = deployment_id) %>% 
  mutate(sonobuoy_number = as.numeric(sonobuoy_number))%>%
  filter(`recording_quality_code` != "UNUSABLE")%>%
  filter(!(cruise == 1621 & sonobuoy_number == 7))
saveRDS(sb_review, file = here("data", "be_review.rds"))


### sb_raven 
# Validate and then merged selection tables
myPath <- here("brydesData/")# 

validation_results <- validate_effort_data(myPath)

# Optionally save results to review
if (validation_results$has_errors) {
  write_csv(validation_results$errors, "validation_errors.csv")
  write_csv(validation_results$summary, "validation_summary.csv")
}

be_raven <- merge_raven_tables(folder_path = myPath) 

saveRDS(be_raven, file = here("data", "be_raven.rds"))

### sb_effort
# process Raven Selection Tables for Effort Analysis
myPath <- here("brydesData/")# 
be_effort <- process_effort_analysis(myPath)
    
# Add sonobuoy_number column 
be_effort <- be_effort %>%
  dplyr::mutate(sonobuoy_number = as.numeric(str_remove(recording_name, "SBDAT"))) 
saveRDS(be_effort, file = here("data", "be_effort.rds"))

### sb_deploy 
# ALL the deployments, then I filter them by the buoys that were reviewed
be_deploy <- read_csv(here("data/surveyData.csv")) %>%
  clean_names() %>%
  dplyr::mutate(
    cruise = as.numeric(cruise),
    sonobuoy_number = as.numeric(sonobuoy_number)
  ) %>% 
  dplyr::semi_join(
    sb_review %>% dplyr::mutate(
      cruise = as.numeric(cruise),
      sonobuoy_number = as.numeric(sonobuoy_number)
    ), 
    by = c("cruise", "sonobuoy_number")
  ) 

## merge and validate deployments
brydes_df <- merge_and_validate_deployments(be_deploy, be_effort, be_review)

saveRDS(brydes_df, file = here("data", "brydes_df.rds"))

```


# Brydes Map

```{r map}
source(here::here("_code/detection_map.R"))

detection_df <- readRDS(here("data/brydes_df.rds"))
output_name <- "brydes_detection_map"

detection_map(detection_df, output_name, point_size = "auto",
              padding_deg = 1, dpi = 300, fig_width = 10,
              fig_height = 8, bathy_res = 0.25, keep_bathy = TRUE,
               xlim = c(-180, -70), ylim = c(-10, 50))
```

# Table of deployments, good recording time, species deployed on, species within 10nmi or 2 hours

Cruise name, deploy ID, date/time, Sighting # & Species, Species w/in 10nmi or 2 hours
```{r deployTable}



```

 
#Table of detections
For each call type, give median and 10th, 90th percentiles for: duration of call (s), low frequency (Hz), high frequency (Hz), center frequency (Hz), peak frequency (Hz), delta frequency (Hz), and signal-to-noise ratio (dB). Where appropriate, give ICI and IPI.





