---
title: "brydes"
output: html_document
---
```{r prep}
source(here::here("_code/_commonR.R"))

surveyData <- read_csv(here("_data/surveyData.csv")) %>%
  clean_names()

review <- read.csv(here("_data/swfsc_sonobuoy_manualReview.csv"))%>%
  clean_names() 
#Next, merge these so that all the files in review have the important info (date, time, lat long,etc). Might need to be sure there is a specific ID column like CruiseNum_SBnum or something

```
# Brydes Map

Map with bathymetry 
- Small black dots for brydes whale detections 
- Small yellow dots for brydes whale detections with calls
- Small yellow stars for brydes whale detections with 500 Hz pulses 

# Table of deployments, good recording time, species deployed on, species within 10nmi or 2 hours

Cruise name, deploy ID, date/time, Sighting # & Species, Species w/in 10nmi or 2 hours
```{r deployTable}



```

 
#Table of detections
For each call type, give median and 10th, 90th percentiles for: duration of call (s), low frequency (Hz), high frequency (Hz), center frequency (Hz), peak frequency (Hz), delta frequency (Hz), and signal-to-noise ratio (dB). Where appropriate, give ICI and IPI.